---
title: "Data Exploration Assignment"
author: "Mohammed Ismail"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


####RESEARCH QUESTION:  
####The College Scorecard was released at the start of September 2015.  Among colleges that predominantly grant bachelors degrees, did the release of the Scorecard shift student interest to high-earnings colleges relative to low-earnings ones (as proxied by Google searches for keywords associated with those colleges)?

###Libraries


```{r}
library(tidyverse)
library(readr)
library(readxl)
library(fixest)
library(purrr)
```


##Pulling the data 
```{r}
id_name_link <- read_csv("../Rawdata/id_name_link.csv")
Scorecard <- read_csv("../Rawdata/Most+Recent+Cohorts+(Scorecard+Elements).csv")
Trends_files <- list.files(path = 'Rawdata', pattern = 'trends_up_to_',full.names = TRUE)
print(Trends_files)
Trends_map=map_df(Trends_files,~read_csv(.))
print(Trends_map)
Trends <- rbind(Trends_map)
summary(Trends)
```

##Initial cleaning to remove NULLS
```{r}
Trends <- Trends %>% 
  na.omit()
summary(Trends)
```

##Playing with Data and exploring
##Ignore(Ended up not using any of it)
```{r}
#id_Scorecard <- merge(x = id_name_link, y = Scorecard, by = c('UNITID', 'OPEID'), all.x = TRUE)
#scorecard_all <- merge(x = ID_Scorecard, y = trends, by = 'schname', all.x = TRUE)
#scorecard_all <- scorecard_all %>% select(-INSTNM)
#scorecard_all <- scorecard_all %>% na.omit()
#Data_To_Play_With <- scorecard_all %>%rename(med_earn = 'md_earn_wne_p10-REPORTED-EARNINGS') %>%filter(PREDDEG == 3) %>%
#  filter(med_earn != 'NULL') %>%filter(med_earn != 'PrivacySuppressed') %>%mutate(med_earn = as.numeric(med_earn))
#Data_To_Play_With <- Data_To_Play_With %>% select('UNITID', 'OPEID', 'schname', 'PREDDEG', 'keyword', 'monthorweek', 'keynum', 'index',
#        'med_earn', 'CONTROL')
#median_earnings <- median(Data_To_Play_With$med_earn)
##The data has a median of 41,800 which could be used to filter the high and low earnings.
```

##Separating by month and year to fit format
```{r}
Trends_month_or_week <- Trends %>%mutate(Y = substr(monthorweek,1,4)) %>%mutate(M = substr(monthorweek,6,7)) %>%
  mutate(MM_YYYY = paste0(M, sep = '-', Y)) %>%select(1,2,6,9)


Trends_month_or_week <- Trends_month_or_week %>% group_by(schid, schname, MM_YYYY) %>% summarise(Ind_sum = sum(index))
```

##Removing duplicate names
```{r}
unique_schname <- id_name_link %>% group_by(schname) %>% filter(n() == 1)

unique_schname <- unique_schname %>% mutate(merge_col = paste0(unitid, sep = ':', opeid))

Scorecard <- Scorecard %>% mutate(merge_col = paste0(UNITID, sep = ':', OPEID))
```

##Merging the variables and filtering for schools for median income
```{r}
score_schname <- merge(Scorecard, unique_schname, by = "merge_col", all = TRUE)

score_schname_red <- score_schname %>% select(126,5,122,11) %>% filter(PREDDEG == 3) %>% 
  filter(`md_earn_wne_p10-REPORTED-EARNINGS` != 'NULL' & `md_earn_wne_p10-REPORTED-EARNINGS` != 'PrivacySuppressed')

score_schname_red$md_income <- as.numeric(score_schname_red$`md_earn_wne_p10-REPORTED-EARNINGS`)
```

##Scorecard income is grouped based if less than 35000 to LOW
##and if above 75000 to HIGH
##between the values to MID.
```{r}
score_schname_red <- score_schname_red %>% mutate(Income_Bracket = case_when(md_income < 35000 ~ 'Low',md_income > 75000 ~ 'High',
                                                                             TRUE ~ 'Mid')) %>% select(1,2,5,6)
```


##Meriging Trend with filtered data
##Setting variables for if pre or post scorecard
```{r}
final_merge <- merge(Trends_month_or_week, score_schname_red, by.x = "schname", by.y = "schname", all.x = TRUE)
final_df <- final_merge %>% na.omit()
final_df$MM_YYYY <- as.Date(paste0(substr(final_df$MM_YYYY,1,2), sep = '/', '01', sep = '/', substr(final_df$MM_YYYY,4,7)), format = '%m/%d/%Y')
final_df <- final_df %>% mutate(norm_ind = (Ind_sum - mean(Ind_sum))/sd(Ind_sum))
final_df <- final_df %>% mutate(Timing = case_when(MM_YYYY < '2015-09-01' ~ 'Pre', MM_YYYY >= '2015-09-01'~ 'Post')) %>% 
  mutate(High = case_when(Income_Bracket == 'High' ~ 1, TRUE ~ 0)) %>% 
  mutate(Mid = case_when(Income_Bracket == 'Mid' ~ 1, TRUE ~ 0)) %>% 
  mutate(Low = case_when(Income_Bracket == 'Low' ~ 1, TRUE ~ 0)) %>% 
  mutate(Pre = case_when(MM_YYYY < '2015-09-01' ~ 1, TRUE ~ 0)) %>% 
  mutate(Post = case_when(MM_YYYY > '2015-09-01' ~ 1, TRUE ~ 0))
```

##Filtering out pre and post dates.
```{r}
final_df_pre_sc <- final_df %>% filter(MM_YYYY < '2015-09-01')
final_df_post_sc <- final_df %>% filter(MM_YYYY > '2015-09-01')
```

##Data regression
```{r}
m1 <- feols(norm_ind ~ Income_Bracket + Timing + Income_Bracket*Timing, data = final_df %>% filter(Income_Bracket != 'Mid'))
etable(m1)

m2 <- feols(norm_ind ~ Current_Treatment | Income_Bracket + Timing, data = final_df %>% filter(Income_Bracket != 'Mid') %>% mutate(Current_Treatment = Income_Bracket == 'High' & Timing == 'Post'))
etable(m2)
```

##Final ANALYSIS
```{r}
did <- final_df %>% filter(Income_Bracket != 'Mid') %>% group_by(Income_Bracket, Timing) %>% summarise(mean = mean(norm_ind))

did_effect <- (did$mean[did$Income_Bracket == 'High' & did$Timing == 'Post'] - did$mean[did$Income_Bracket == 'High' & did$Timing == 'Pre']) - (did$mean[did$Income_Bracket == 'Low' & did$Timing == 'Post'] - did$mean[did$Income_Bracket == 'Low' & did$Timing == 'Pre'])

round(did_effect, 4)
```


##Comparing for pre and post scorecard
```{r}
#Box plot
bp<-ggplot(final_df %>% filter(Income_Bracket != 'Mid'), aes(y = norm_ind, x = Income_Bracket, color = Timing)) + geom_boxplot()
bp + scale_fill_manual(breaks = c("2", "1", "0.5"), 
                       values=c("red", "blue", "green"))


#Scatter plot
sp<-ggplot(final_df %>% filter(Income_Bracket != 'Mid'), aes(y = norm_ind, x = MM_YYYY, color = Income_Bracket, shape = Timing))+geom_smooth(method = 'lm')
sp + scale_color_manual(breaks = c("8", "6", "4"),
                        values=c("red", "blue", "green"))
```
